---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PoiMA
Jiahui Zou

<!-- badges: start -->
<!-- badges: end -->

Model averaging is a popular methodology to promote prediction and deal with the uncertainty of covariate choice when we use a statistical model. This package, PoiMA, is designed to implement model averaging for Poisson regressions, in which the criterion to calculate model-averaging weight is unbiased.
 
## Model
Suppose we observe a dataset $\{(y_i,\mathbf{x}_i),i=1,\ldots,n\}$, where $y_i$ is the realization of a count variable $Y_i$ measuring the number of occurrences of an even in a given interval for the $i$th individual, $\mathbf{x}_i$ is a $p-$dimensional vector, and $n$ denotes the number of observations. A Poisson distribution is commonly used, which associates the probability of $Y_i$ events with $\mathbf{x}_i$ as follows:
$$
\begin{align}
P(Y_i=y_i|\mu_i)=\frac{\exp(-\mu_i)\mu_i^{y_i}}{y_i!},
\end{align}
$$
where $\mu_i$ is the Poisson incidence rate that depends on $\mathbf{x}_i$ as $\mu_i=\exp(\mathbf{x}_i^\top\boldsymbol{\beta})$ and $\boldsymbol{\beta}$ is the associated $p\times 1$ coefficient vector.


In practice, not all covariates are ``useful'' in predicting $y_i$, and researchers are typically uncertain regarding which should be included in the model \emph{ex ante}. Hence, a number of candidate models with different specifications of $\mathbf{x}_i$ are considered.
Let $S$ be the number of candidate models and let $\Pi_s$ be a $p\times p_s$ selection matrix that consists of 0's and 1's and selects $p_s$ covariates ($p_s\le p$) for the $s^{th}$ model. Denote the $p_s$-dimensional covariate vector in the $s$th model by $\mathbf{x}^\top_{(s),i}=\mathbf{x}_i^\top\Pi_s$. We model the Poisson incidence rate of this candidate model as
$$
\begin{align}
\mu_{(s),i}=\exp(\mathbf{x}_{(s),i}^\top\boldsymbol{\beta}_{(s)}).
\end{align}
$$
After solving each candidate model, we need to unify the dimension of the coefficient estimates of each candidate model by $\hat{\boldsymbol{\beta}}_s=\Pi_s\hat{\boldsymbol{\beta}}_{(s)}$.
Then, we can compute the model averaging estimates of coefficient $\boldsymbol{\beta}$ by
$$
\begin{align}
\hat{\boldsymbol{\beta}}(\mathbf{w})=\sum_{s=1}^S {w}_s\hat{\boldsymbol{\beta}}_s,
\end{align}
$$
where $\mathbf{w}=(w_1,w_2,...,w_S)^\top$ is the weight vector belonging to the set
$\mathcal{W}=\{\mathbf{w}\in[0,1]^S:\sum_{s=1}^S w_s=1\}$.

The model-averaging weight $\mathbf{w}$ is calculated by minimizing the following criterion,
$$
\begin{align}
C(\mathbf{w})=\log f(\mathbf{y})+\sum_{i=1}^n\left[\hat{\boldsymbol{\mu}}_i(\mathbf{w},\mathbf{y})+\log(\mathbf{y_i!})-y_i\log\{\hat{\boldsymbol{\mu}}_i(\mathbf{w},\mathbf{y}^{(y_i-1)})\}
\right],
\end{align}
$$
where $\mathbf{y}^{(y_i-1)}=(y_1,\ldots, y_{i-1},y_i-1,y_{i+1},\ldots,y_n)\top$, which replace the $i$th element of $\mathbf{y}$ by $y_i-1$, $\hat{\boldsymbol{\mu}}(\mathbf{w},\mathbf{y}^{(y_i-1)})=\left(\hat{\mu}_1(\mathbf{w},\mathbf{y}^{(y_i-1)}),\ldots,\hat{\mu}_n(\mathbf{w},\mathbf{y}^{(y_i-1)}\right)\top$. Specially, when $y_i=0$, $y_i\log\{\hat{\mu}_i(\mathbf{w},\mathbf{y}^{(y_i-1)})\}=0$.



## Installation

You can install the development version of PoiMA from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("zoujiahuibin/PoissonMA")
```

## Example

There are basic examples which show you how to solve a common problem:


```{r example1}
options(warn =-1)
library(PoiMA)
library(mvtnorm)
library(magrittr)
beta0=rep(0.1,5)
X=rmvnorm(100,mean=rep(0,5))
lambda=exp(X%*%beta0)
y=rpois(100,lambda)
ModelSetup(y,X,'nested',intercept=FALSE)
```

```{r example2}
options(warn =-1)
library(PoiMA)
library(mvtnorm)
library(magrittr)
beta0=rep(0.1,5)
X=cbind(1,rmvnorm(100,mean=rep(0,4)))
lambda=exp(X%*%beta0)
y=rpois(100,lambda)
ModelSetup(y,X,'nested',intercept=TRUE)
```
### Example 2

```{r example3}
options(warn =-1)
library(PoiMA)
library(mvtnorm)
library(magrittr)
library(Rsolnp)
beta0=rep(0.1,5)
X=rmvnorm(100,mean=rep(0,5))
lambda=exp(X%*%beta0)
y=rpois(100,lambda)
PoissonMA(y,X, modeltype='nested',intercept=FALSE)
```



```{r example4}
options(warn =-1)
library(PoiMA)
library(mvtnorm)
library(magrittr)
library(Rsolnp)
beta0=rep(0.1,5)
X=cbind(1,rmvnorm(100,mean=rep(0,4)))
lambda=exp(X%*%beta0)
y=rpois(100,lambda)
PoissonMA(y,X, modeltype='nested',intercept=TRUE)
```



# Reference

Zou, J., Wang, W., Zhang, X. and Zou, G. (2022). Optimal model averaging for divergent-dimensional Poisson regressions. Econometric Reviews 41(7): 775--805.
